---
title: "Data Analysis For Motor Trend Car Road Tests"
author: "Devneet Mohanty"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
format:
  html:
    code-fold: true
    code-tools: true
    code-line-numbers: true
    fontsize: 1.2em
    toc: true
    toc-title: '**Contents**'
    embed-resources: true
    grid:
      body-width: 800px
    theme: 
      - superhero
      - custom-theme.css
    css: custom-theme.css
  revealjs: 
    slide-number: true
    incremental: false
    auto-stretch: false
    embed-resources: true
    center: true
    theme:
       - custom-theme.css
urlcolor: blue
progress: true
controls: true
execute: 
  eval: true
  echo: false
  warning: false
---

## Background

The **`mtcars`** data originated in the **1974 Motor Trend** magazine and records fuel economy alongside various mechanical specs for 32 different car models. In this brief analysis, I will:

1.  Calculate basic summary measures for **MPG** and **HP**.  
2.  Create an interactive scatter showing how miles per gallon changes with engine power.  
3.  Offer a few observations on what the plot reveals.

## Summary Statistics

First, we will compute means and standard deviations for our two key metrics using `dplyr` package.

```{r}
#| label: desc-stat

# Importing the `dplyr` library:
library(dplyr)

# Viewing the summary of the `mtcars` dataset:
stats <- mtcars %>%
  summarise(
    Mean_MPG = mean(mpg),
    SD_MPG   = sd(mpg),
    Mean_HP  = mean(hp),
    SD_HP    = sd(hp)
  )

stats
```

## Scatterplot Between Miles Per Gallon & Horsepower

Below is an interactive chart built using `ggplot2` and `plotly` packages. We’ve added a dashed linear trend and shaded 95% confidence band to highlight the overall relationship between [Miles Per Gallon (mpg)]{style="color: teal;"} and [Horsepower (hp)]{style="color: orchid;"}.

```{r}
#| label: mpg-vs-hp

# Importing the libraries for plotting:
library(ggplot2)
library(plotly)

# Creating the plotting object:
p <- ggplot(
  mtcars, 
  aes(x = hp, y = mpg)
  ) +
  # Setting points for the plot:
  geom_point(size = 3, color ='purple') +
  # Setting regression line for the plot:
  geom_smooth(
    method    = "lm",
    se        = TRUE,
    linetype  = "dashed",
    alpha     = 0.2
  ) +
  # Setting the labels, title and text for the plot:
  labs(
    title = "Fuel Efficiency vs. Engine Power",
    x     = "Horsepower (HP)",
    y     = "Miles per Gallon (MPG)",
    shape = "Transmission"
  ) +
  theme_minimal(base_size = 14)

# Creating the interactive plot:
ggplotly(p)
```

## Interpretation Of Relationship

- There’s a clear inverse trend: as horsepower goes up, miles per gallon tends to go down.

- The shaded ribbon around the dashed line shows our 95% confidence interval for the fitted regression.

- Understanding this trade-off can guide decisions about performance versus economy when selecting a vehicle.