---
title: "Effects of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Devneet Mohanty"
format:
  html:
    code-fold: true
    code-tools: true
    code-line-numbers: true
    fontsize: 1.2em
    toc: true
    toc-title: '**Contents**'
    embed-resources: true
    grid:
      body-width: 800px
    theme: 
      - superhero
      - custom-theme.css
    css: custom-theme.css
  revealjs: 
    slide-number: true
    incremental: false
    auto-stretch: false
    embed-resources: true
    center: true
    theme:
      - custom-theme.css
urlcolor: blue
progress: true
controls: true
execute: 
  eval: true
  echo: false
  warning: false
---

## Background

A classic biological experiment was conducted to determine the effects of **vitamin C** on the tooth length of guinea pigs. The data records the **length of odontoblasts (cells responsible for tooth growth)** under different supplement types and dosages in 60 guinea pigs. In this brief analysis, I will:

1. Compute summary statistics for **tooth length**, categorized by **supplement type** and **dose level**.  
2. Visualize the impact of **vitamin C dosage** and **delivery method** on guinea pig tooth growth using an interactive boxplot.  
3. Interpret whether the **form of supplementation** (OJ vs. VC) and **dosage amount** significantly influence tooth development.  

---

## Summary Statistics

We begin by calculating the mean and standard deviation of **tooth length**, grouped by **supplement type** and **dosage level**, using the `dplyr` package.

```{r}
#| label: desc-stat-tooth
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(knitr)
library(kableExtra)

# Calculate summary statistics
stats_tooth <- ToothGrowth %>%
  group_by(Supplement = supp, Dose_mg = dose) %>%
  summarise(
    Mean_Tooth_Length = round(mean(len), 2),
    SD_Tooth_Length   = round(sd(len), 2),
    .groups = "drop"
  )

# Display as styled HTML table
stats_tooth %>%
  kable("html", caption = "Summary Statistics of Tooth Length by Supplement and Dose") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  )
```

## Effect of Supplement and Dose on Tooth Length

The following interactive boxplot compares [Tooth Length (len)]{style="color: darkgreen;"} across different dosage levels and supplement types. It helps us observe how increasing the vitamin C dose affects tooth growth and whether the delivery method (Orange Juice vs. Vitamin C supplement) plays a role.

This visualization allows us to:

- Observe how increasing doses of Vitamin C (0.5, 1.0, and 2.0 mg) influence the average tooth growth in guinea pigs.

- Examine whether the form of Vitamin C delivery — either as Orange Juice (OJ) or a pure supplement (VC) — has a differential effect on tooth development.

- Identify potential interaction effects between dosage and supplement type, especially if the growth trends differ significantly across groups.

- By comparing the spread and central tendency (median) of tooth length distributions within each group, we aim to determine the optimal dose and delivery method for promoting tooth growth.

```{r}
#| label: toothgrowth-boxplot
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(plotly)

# Convert dose to factor for plotting
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Create boxplot
p_box <- ggplot(ToothGrowth, aes(x = dose, y = len, fill = supp)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red", outlier.size = 2) +
  labs(
    title = "Tooth Length by Dose and Supplement Type",
    x = "Supplement Type Dose (mg)",
    y = "Tooth Length (mm)",
    fill = "Supplement Type"
  ) +
  theme_minimal(base_size = 14)

ggplotly(p_box)
```


## Interpretation of Relationship

- There’s a clear positive effect of increasing vitamin C dosage on tooth length: higher doses generally result in longer teeth.

- At each dose level, orange juice (OJ) tends to produce slightly greater tooth growth than the synthetic vitamin C supplement (VC), especially at the 0.5 mg and 1.0 mg levels.

- These patterns suggest that both dosage and delivery method significantly influence tooth development in guinea pigs, with a potential interaction between the two factors.